### 2008-05-30 doc/sec/security_infrastructure.tex
Nice formating

{\raggedleft NORDUGRID-XXXXXXX-NN

@@

\chapter{Introduction}

\label{sec:intro}

@@

\chapter{Security architecture in HED: SecHandlers and PDP}

\chapter{Policy Evaluation engine}

\section{ArcPDP Policy Evaluator}

\chapter{Secuirty Attributes}

\section{Infrastructure}

Security Attributes represent security related information inside

HED framework and store information representing various aspects

needed to perform authorization decison - identity of client, requested

action, targeted resource, constraint policies. Each kind of Security

Attribute is represented by own class inherited from parent SecAttr
class.

Each Security Attribute stores it's information in internal format
and

provides it to external code in one of predefined formats. Currently

only supported format is Arc Policy/Request.

Collectors of Security Attributes instantiate corresponding classes
and

link them to Secuirity Attributes containers - MessageAuth and

MessageAuthContext representing attributes per request and per session

correspondingly. Current implementations of Security Attributes Collectors

are either integratd into existing MCCs or implemented as separate
SecHandler

plugins. See Appendix A for available collectors and corresponding
Security

Attributes.

Note for service developers: Services may implement own authorization

algorithms. But they may use Security Atributes as well by providing

instances of classes inherited from SecAttr and running them through

either configured or hardcoded processors/PDPs (see below).

Processors of Security Attributes are implemented as Policy Decision
Point

components. Currently there are 2 PDP components available:

{*} Arc PDP makes use of Security Attributes containing identities
of client,

resource and requested action. It evaluates either all or selected
set

of attributes against specified Policy documents thus making it possible

to enforce policies defined/selected by service providers.

{*} Delegation PDP is similar to Arc PDP except that it takes it's
Policy

documents directly from Security Attributes. Differently from Arc
PDP it

is meant to be used for enforcing policies defined by client.

Typical service example - TODO

\section{Available collectors}

{*} TCP

@@

{*} TLS

Information is collected inside TLS MCC. Security Attribute exports

SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/ca
Subject of signer of first certificate in client's chain

SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/chain
Subject of certificate in client's chain - multiple items

SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/subject
Subject of last certificate in client's chain

SubjectAttribute http://www.nordugrid.org/schemas/policy-arc/types/tls/identity
Subject of last non-proxy certificate in client's chain

{*} HTTP

Information is collected inside TLS MCC. Security Attribute exports

Resource http://www.nordugrid.org/schemas/policy-arc/types/http/path
HTTP path without host and port part

Action http://www.nordugrid.org/schemas/policy-arc/types/http/method
HTTP method

{*} SOAP

Information is collected inside TLS MCC. Security Attribute exports

Resource http://www.nordugrid.org/schemas/policy-arc/types/soap/endpoint
WS-Addr To element

Action http://www.nordugrid.org/schemas/policy-arc/types/soap/operation
SOAP top level element name without namespace prefix

Context http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace
Namespace of SOAP top level element

{*} Delegation

This Security Attribute is collected by dedicated Security Handler
plugin.

It extracts policy document stored inside X509 certificate proxy extension

as defined in RFC3820. All certificates in a chain provided by client
are

examined. Proxy certificates with id-ppl-inheritAll property are passed

through and no policy is generated for them. From those with id-ppl-anyLanguage

string content of ProxyPolicy element is extracted and converted into
XML

document. Then document is check to be of Arc Policy kind. If policy
is

not recognized as Arc Policy procedure fails and that causes failure
of

comunication. Proxies with other type of policies including id-ppl-independent

are not accepted and generate immediate failure.

\chapter{Delegation restrictions}

\section{Delegation architecture}

\section{Delegation PDP}

\chapter*{Schemas}

@@

>>>>

{\raggedleft NORDUGRID-TECH-16

@@

\chapter{Introduction\label{sec:introduction}}

@@

\chapter{Security architecture in HED: SecHandlers and PDP\label{sec:architecture}}

\chapter{Policy Evaluation engine\label{sec:policy_engine}}

\section{ArcPDP Policy Evaluator\label{sec:policy_evaluator}}

\chapter{Secuirty Attributes\label{sec:sec_attr}}

\section{Infrastructure\label{sec:sec_attr_infrastructure}}

Security Attributes represent security related information inside
HED framework and store information representing various aspects needed
to perform authorization decison - identity of client, requested action,
targeted resource, constraint policies.

Each kind of Security Attribute is represented by own class inherited
from parent SecAttr class <arc/message/SecAttr.h>. Each Security Attribute
stores it's information in internal format and is capable to export
it to one of predefined formats using Export() method. Currently only
supported format is Arc Policy/Request XML document.

Collectors of Security Attributes instantiate corresponding classes
and link them to Secuirity Attributes containers - MessageAuth <arc/message/MessageAuth.h>
and MessageAuthContext <arc/message/Message.h> storing collected attributes
per request and per session correspondingly. Current implementations
of Security Attributes Collectors are either integrated into existing
MCCs or implemented as separate SecHandler plugins. See Section \ref{sec:sec_attr_collectors}
for available Collectors and corresponding Security Attributes.

\begin{framed}Note for service developers: Services may implement
own authorization algorithms. But they may use Security Atributes
as well by providing instances of classes inherited from SecAttr and
running them through either configured or hardcoded processors/PDPs.\end{framed}

Processors of Security Attributes are implemented as Policy Decision
Point components. Currently there are 2 PDP components available:

\begin{itemize}
\item ArcPDP makes use of Security Attributes containing identities of client,
resource and requested action. It evaluates either all or selected
set of attributes against specified Policy documents thus making it
possible to enforce policies defined/selected by service providers. 
\item (to move to Delegation section) DelegationPDP is similar to ArcPDP
except that it takes it's Policy documents directly from Security
Attributes. Differently from Arc PDP it is meant to be used for enforcing
policies defined by client.
\end{itemize}

\section{Available collectors\label{sec:sec_attr_collectors}}

\subsection{TCP}

@@

\subsection{TLS}

Information is collected inside TLS MCC. Generated Security Attribute
class exports ARC Request with following attributes:

\begin{tabular}{|c|c|c|}
\hline 
\multicolumn{1}{|c|}{Element}&
AttributeId&
Content\tabularnewline
\hline
\hline 
SubjectAttribute&
http://www.nordugrid.org/schemas/policy-arc/types/tls/ca&
Subject of signer of first certificate in client's chain\tabularnewline
\hline 
SubjectAttribute&
http://www.nordugrid.org/schemas/policy-arc/types/tls/chain&
Subject of certificate in client's chain - multiple items\tabularnewline
\hline 
SubjectAttribute&
http://www.nordugrid.org/schemas/policy-arc/types/tls/subject&
Subject of last certificate in client's chain\tabularnewline
\hline 
SubjectAttribute&
http://www.nordugrid.org/schemas/policy-arc/types/tls/identity&
Subject of last non-proxy certificate in client's chain\tabularnewline
\hline
\end{tabular}

\subsection{HTTP}

Information is collected inside TLS MCC. Security Attribute exports
ARC Request with following attributes:

\begin{tabular}{|c|c|c|}
\hline 
Element&
AttributeId&
Content\tabularnewline
\hline
\hline 
Resource&
http://www.nordugrid.org/schemas/policy-arc/types/http/path&
HTTP path without host and port part\tabularnewline
\hline 
Action&
http://www.nordugrid.org/schemas/policy-arc/types/http/method&
HTTP method\tabularnewline
\hline
\end{tabular}

\subsection{SOAP}

Information is collected inside TLS MCC. Security Attribute exports
ARC Request with following attributes

\begin{tabular}{|c|c|c|}
\hline 
Element&
AttributeId&
Content\tabularnewline
\hline
\hline 
Resource&
http://www.nordugrid.org/schemas/policy-arc/types/soap/endpoint&
To element of WS-Addressing structure\tabularnewline
\hline 
Action&
http://www.nordugrid.org/schemas/policy-arc/types/soap/operation&
SOAP top level element name without namespace prefix\tabularnewline
\hline 
Context&
http://www.nordugrid.org/schemas/policy-arc/types/soap/namespace&
Namespace of SOAP top level element\tabularnewline
\hline
\end{tabular}

\subsection{Delegation}

(to move to Delegation section) This Security Attribute is collected
by dedicated Security Handler plugin named {}``\emph{delegation.pdp}''
avaialble in \emph{arcpdc} plugins' library. It extracts policy document
stored inside X509 certificate proxy extension as defined in RFC3820.
All proxy certificates in a chain provided by client are examined
and all available policies are extracted.

Policies are stored in string content of \emph{ProxyPolicy} element
with \emph{id-ppl-anyLanguage}. Extracted content is converted into
XML document. Then document is checked to be of Arc Policy kind. If
policy is not recognized as Arc Policy procedure fails and that causes
failure of comunication.

Proxy certificates with \emph{id-ppl-inheritAll} property are passed
through and no policy document is generated for them. Proxies with
other type of policies including \emph{id-ppl-independent} are not
accepted and generate immediate failure.

\chapter{Delegation restrictions\label{sec:delegation}}

\section{Delegation architecture\label{sec:delegation_architectire}}

\section{Delegation PDP\label{sec:delegation_pdp}}

\chapter*{Schemas and comments\label{sec:schemas}}

@@

\section*{ArcPDP configuration}

\section*{DelegationPDP configuration}
